Grappa: Scalable Computing on Commodity Hardware \\for Irregular Applications

### 2012-07-24 doc/asplos13/intro.tex
i
fixed what i think were typos in 2nd to last paragraph.

Grappa leverages as much freely available and commodity infrastructure as
possible. We use unmodified Linux for the operating system and an
off-the-shelf user-mode infiniband device driver stack~\cite{OFED}. MPI is
used for process setup and tear down. GASnet~\cite{gasnet} is used as the
underlying mechanism for remote memory reads and writes using active message
invocations. To this commodity hardware and software mix Grappa adds three
main software components: (1) a \emph{lightweight tasking} layer that supports
a context switch in as few as 38ns and distributed global load balancing; (2)
a \emph{distributed shared memory} layer that supports customary operations
such as \emph{read and write as well as remote operations} such as
fetch-and-add~\cite{fetchandadd}; and (3) a \emph{message aggregation} layer
that combines short messages to mitigate the aforementioned problem that
commodity networks are designed to achieve peak bandwidth only on large packet
sizes, yet irregular applications tend to fetch only a handful of bytes at a
time. As we will show later, Grappa can tolerate latencies way beyond that of
the network. Therefore, Grappa can afford to \emph{trade latency for
throughput}: by {\em increasing} latency in key components of the system we
are able to increase our effective random access memory bandwidth (by delaying and aggregating messages), our synchronization bandwidth (by delegating operations
to remote nodes), and our ability to improve load imbalance (work stealing
increases latency).

@@

Our evaluation of Grappa shows that it runs several graph-crunching
applications (classic examples of irregular behavior) very efficiently
on a commodity cluster. Our yardstick for comparison is the XMT hardware
itself. Using the same number of network interfaces, Grappa is in the
same ballpark as the XMT: For unbalanced tree search, Grappa is over 3X
faster and shows greatly improved scalability; conversely, for breadth
first search and betweenness centrality Grappa is 2.5X slower.  In
Section~\ref{sec:results} we explore the factors that underpin this
performance. Most importantly, however, for significantly less real
world cost, users can \emph{add} significantly more processors to a
commodity cluster than an XMT machine and use Grappa to achieve scalable
performance.

>>>>

Grappa leverages as much freely available and commodity infrastructure as
possible. We use unmodified Linux for the operating system and an
off-the-shelf user-mode infiniband device driver stack~\cite{OFED}. MPI is
used for process setup and tear down. GASnet~\cite{gasnet} is used as the
underlying mechanism for remote memory reads and writes using active message
invocations. To this commodity hardware and software mix Grappa adds three
main software components: (1) a \emph{lightweight tasking} layer that supports
a context switch in as few as 38ns and distributed global load balancing; (2)
a \emph{distributed shared memory} layer that supports normal access operations
such as \emph{read} and \emph{write} as well as synchronizing operations such as \emph{fetch-and-add}\cite{fetchandadd}; and (3) a \emph{message aggregation} layer
that combines short messages to mitigate the aforementioned problem that
commodity networks are designed to achieve peak bandwidth only on large packet
sizes, yet irregular applications tend to fetch only a handful of bytes at a
time. As we will show later, Grappa can tolerate latencies way beyond that of
the network. Therefore, Grappa can afford to \emph{trade latency for
throughput}: by {\em increasing} latency in key components of the system we
are able to increase our effective random access memory bandwidth (by delaying and aggregating messages), our synchronization bandwidth (by delegating operations
to remote nodes), and our ability to improve load imbalance (work stealing
increases latency).

@@

Our evaluation of Grappa shows that it runs several graph-crunching
applications (classic examples of irregular behavior) very efficiently
on a commodity cluster. Our yardstick for comparison is the XMT hardware
itself. Using the same number of network interfaces, Grappa is in the
same ballpark as the XMT: For unbalanced tree search, Grappa is over 3X
faster and shows greatly improved scalability; conversely, for breadth
first search and betweenness centrality Grappa is 2.5X slower.  In
Section~\ref{sec:evaluation} we explore the factors that underpin this
performance. Most importantly, however, for significantly less real
world cost, users can \emph{add} significantly more processors to a
commodity cluster than an XMT machine and use Grappa to achieve scalable
performance.
